Silent pessimists: researchers' perceptions of malaria eradication

### 2017-12-27 paper.Rmd
update

---
title: "Silent pessimists: researchers' perceptions of malaria eradication and implications for health spending"
bibliography: library.bib
output: 
  pdf_document:
    latex_engine: xelatex
    includes:
      in_header: preamble.sty
---

@@

@@

cols <- colorRampPalette(c('red', 'orange', 'blue') #brewer.pal(n = 9, 'Spectral')
                         )(length(unique(x$citations)))
x1 <- ggplot(data = x %>% filter(!is.na(citations)),
       aes(x = as.numeric(gsub('years_', '', key)),
           y = value,
           color = citations,
           group = citations)) +
  geom_point(alpha = 0.7) +
  geom_line(alpha = 0.6) +
  theme_paper() +
  scale_color_manual(name = 'Number of citations',
                     values = cols) +
  labs(x = 'Years',
       y = 'Likelihood of eradication',
       subtitle = 'A: Difference in perceived likelihood of eradication by citations')
```

```{r}
x2 <- 
  ggplot(data = heckman,
       aes(x = years,
           y = value,
           color = key,
           lty = key)) +
  geom_line(alpha = 0.8) +
  geom_point(alpha = 0.6) +
  theme_paper() +
  labs(x = 'Years',
       y = 'Likelihood of eradication',
       subtitle = 'B: Perceived likelihood of eradication: Heckman method and unadjusted') +
  scale_linetype_manual(name = '', values = 1:2) +
  scale_color_manual(name = '', values = c('blue', 'darkorange'))
```

```{r, fig.cap = 'Sample selection bias and adjustment', fig.height = 3.5}
Rmisc::multiplot(x1, x2, cols = 2)
```

Of the below two sections only one will remain in the paper: either (i) de-biasing using weighted binomial regression or (ii) de-biasing using Heckman.

## De-biasing using weighted binomial regression

```{r}
# Estimate model for de-biasing sample selection without Heckman.
response_model <- glm(responded ~ bin_citations + sex, data = df, family = binomial('logit'))
# Use model to predict likelihood of response
df$likelihood_of_response <- predict(response_model, df, type = 'response')
# Use inverse of likelihood of response to assign weight for model
df$weight <- 1 - df$likelihood_of_response
```

## De-biasing using Heckman

In order to de-bias sample selection, we employed Heckman's two-step correction method [@Heckman_1979]. Given that sex affected response rate but not response itself, we were able to use sex as an instrumental variable, introducing it in both the selection and response models. Our selection models, which estimated the likelihood of survey response as a probit function of sex allowed us to estimate (via the Inverse Mills ratio) an expected error term for our outcome models, in which we quantify the effect of the number of citations on the perceived average likelihood of eradication at 10, 20, 30, 40, and 50 years (employing a separate model for each time period). Figure 3 shows the difference in the aggregate perceived likelihood of eradication before and after implementing Heckman's two-step model.

Table 2 shows the Heckman model Heckman estimates for all five models.

```{r}
rn <- names(heck_10$coefficients)
heck <- data_frame(Coefficient = rn,
                   `Model 10` = heck_10$coefficients,
                   `Model 20` = heck_20$coefficients,
                   `Model 30` = heck_30$coefficients,
                   `Model 40` = heck_40$coefficients,
                   `Model 50` = heck_50$coefficients)
for (j in 2:ncol(heck)){
  heck[,j] <- round(heck[,j], digits = 3)
}
heck$Coefficient <- gsub('bin_citations', 'Citations ', heck$Coefficient)
heck$Coefficient <- gsub('invMillsRatio', 'Inverse Mills Ratio', heck$Coefficient)
heck$Coefficient <- gsub('sigma', 'Sigma', heck$Coefficient)
heck$Coefficient <- gsub('rho', 'Rho', heck$Coefficient)
heck$Coefficient <- gsub('sexmale', 'Sex male', heck$Coefficient)
heck$Coefficient <- paste0(c(rep('Selection ', 4), rep('Outcome ', 2), rep('', 3)), heck$Coefficient)
knitr::kable(heck,
             caption = 'Heckman selection model coefficients')
```

>>>>

---
title: "Silent pessimists: researchers' perceptions of malaria eradication"
bibliography: library.bib
output: 
  pdf_document:
    latex_engine: xelatex
    includes:
      in_header: preamble.sty
---

@@

\newpage

\noindent\fbox{%
    \parbox{\textwidth}{%
        Of the below two columns only one will remain in the paper: either (i) de-biasing using weighted binomial regression or (ii) de-biasing using Heckman.
    }%
}

\begin{multicols}{2}

\subsection*{De-biasing using weighted binomial regression}

In order to de-bias sample selection, we estimated a binomial logistic regression model for the likelihood of response as a function of sex and (binned) number of citations, and then used the inverse of the selection model's predictions as weights in a simple linear model to adjust our estimates. We run weighted model to estimate the likelihood of eradication at 10, 20, 30, 40, and 50 years (employing a separate model for each time period). Figure 3 shows the difference in the aggregate perceived likelihood of eradication before and after adjusting for sample selection.

```{r}
# Estimate model for de-biasing sample selection without Heckman.
response_model <- glm(responded ~ bin_citations + sex, data = df, family = binomial('logit'))
# Use model to predict likelihood of response
df$likelihood_of_response <- predict(response_model, df, type = 'response')
# Use inverse of likelihood of response to assign weight for model
df$weight <- 1 - df$likelihood_of_response
# Run likelihood of eradication models
for (i in seq(10, 50, 10)){
  assign(paste0('formula_simple_years_', i),
         as.formula(paste0('years_',
                           i,
                           '_numeric ~ sex + bin_citations')))
  assign(paste0('simple_', i),
         lm(get(paste0('formula_simple_years_', i)), data = df))
  assign(paste0('simple_', i, '_weighted'),
         lm(get(paste0('formula_simple_years_', i)), data = df,
            weights = df$weight))

}

# Get aggregate adjusted vs unadjusted likelihoods of eradications
simple <- data_frame(years = seq(10, 50, 10),
                     adjusted = NA,
                     unadjusted = NA)
for (i in seq(10, 50, 10)){
   simple$unadjusted[simple$years == i] <- mean(df[,paste0('years_', i, '_numeric')], na.rm = TRUE)
   simple$adjusted[simple$years == i] <- weighted.mean(x = df[,paste0('years_', i, '_numeric')],
                                                       w = ifelse(is.na(df$weight), 0,
                                                                  df$weight), na.rm = TRUE)
  # simple_model <- get(paste0('simple_', i))
  # weighted_model <- get(paste0('simple_', i, '_weighted'))

}
simple <- gather(simple, key, value, adjusted:unadjusted)
g2 <- ggplot(data = simple,
       aes(x = years,
           y = value,
           color = key,
           lty = key)) +
    geom_line(alpha = 0.8) +
  geom_point(alpha = 0.6) +
  theme_paper() +
  labs(x = 'Years',
       y = 'Likelihood of eradication',
       subtitle = 'B: Perceived likelihood of eradication: Adjusted and unadjusted') +
  scale_linetype_manual(name = '', values = 1:2) +
  scale_color_manual(name = '', values = c('blue', 'darkorange'))
x <- df %>%
  group_by(citations = bin_citations) %>%
  summarise(years_10 = mean(years_10_numeric, na.rm = TRUE),
            years_20 = mean(years_20_numeric, na.rm = TRUE),
            years_30 = mean(years_30_numeric, na.rm = TRUE),
            years_40 = mean(years_40_numeric, na.rm = TRUE),
            years_50 = mean(years_50_numeric, na.rm = TRUE),
            n = length(unique(email))) %>%
  ungroup %>%
  gather(key, value, years_10:years_50) %>%
  filter(!is.na(citations))

cols <- colorRampPalette(c('red', 'orange', 'blue') #brewer.pal(n = 9, 'Spectral')
                         )(length(unique(x$citations)))
g1 <- ggplot(data = x %>% filter(!is.na(citations)),
       aes(x = as.numeric(gsub('years_', '', key)),
           y = value,
           color = citations,
           group = citations)) +
  geom_point(alpha = 0.7) +
  geom_line(alpha = 0.6) +
  theme_paper() +
  scale_color_manual(name = 'Number of citations',
                     values = cols) +
  labs(x = 'Years',
       y = 'Likelihood of eradication',
       subtitle = 'A: Difference in perceived likelihood of eradication by citations')
```

```{r, fig.height = 7.5}

Rmisc::multiplot(g1, g2, cols = 1)

# # Combine into a dataframe
# simple_df <- expand.grid(bin_citations = sort(unique(df$bin_citations)),
#                          sex = sort(unique(df$sex))) %>%
#   mutate(simple = NA,
#          weighted = NA)
# out_list <- list()
# counter <- 0
# for(i in seq(10, 50, 10)){
#   counter <- counter + 1
#   this_df <- simple_df %>%
#     mutate(years = i)
#   simple_model <- get(paste0('simple_', i))
#   weighted_model <- get(paste0('simple_', i, '_weighted'))
#   this_df$simple <- predict(simple_model, this_df)
#   this_df$weighted <- predict(weighted_model, this_df)
#   out_list[[counter]] <- this_df
# }
# simple_df <- bind_rows(out_list)
# simple_df <- simple_df %>%
#   gather(key, value, simple:weighted)
# # Just aggregate sexes
# simple_df <- simple_df %>%
#   group_by(bin_citations, years, key) %>%
#   summarise(value = mean(value)) %>%
#   ungroup
# 
# ggplot(data = simple_df,
#        aes(x = years,
#            y = value,
#            color = bin_citations)) +
#   geom_line()
```

\vfill\null
\columnbreak

\subsection*{De-biasing using Heckman}

In order to de-bias sample selection, we employed Heckman's two-step correction method. Given that sex affected response rate but not response itself, we were able to use sex as an instrumental variable, introducing it in both the selection and response models. Our selection models, which estimated the likelihood of survey response as a probit function of sex allowed us to estimate (via the Inverse Mills ratio) an expected error term for our outcome models, in which we quantify the effect of the number of citations on the perceived average likelihood of eradication at 10, 20, 30, 40, and 50 years (employing a separate model for each time period). Figure 3 shows the difference in the aggregate perceived likelihood of eradication before and after implementing Heckman's two-step model.

@@

cols <- colorRampPalette(c('red', 'orange', 'blue') #brewer.pal(n = 9, 'Spectral')
                         )(length(unique(x$citations)))
x1 <- ggplot(data = x %>% filter(!is.na(citations)),
       aes(x = as.numeric(gsub('years_', '', key)),
           y = value,
           color = citations,
           group = citations)) +
  geom_point(alpha = 0.7) +
  geom_line(alpha = 0.6) +
  theme_paper() +
  scale_color_manual(name = 'Number of citations',
                     values = cols) +
  labs(x = 'Years',
       y = 'Likelihood of eradication',
       subtitle = 'A: Difference in perceived likelihood of eradication by citations')

x2 <- 
  ggplot(data = heckman,
       aes(x = years,
           y = value,
           color = key,
           lty = key)) +
  geom_line(alpha = 0.8) +
  geom_point(alpha = 0.6) +
  theme_paper() +
  labs(x = 'Years',
       y = 'Likelihood of eradication',
       subtitle = 'B: Perceived likelihood of eradication: Heckman method and unadjusted') +
  scale_linetype_manual(name = '', values = 1:2) +
  scale_color_manual(name = '', values = c('blue', 'darkorange'))
```

```{r fig.height = 7.5}
Rmisc::multiplot(x1, x2, cols = 1)
```

```{r, eval = FALSE}
# Table 2 shows the Heckman model Heckman estimates for all five models.

rn <- names(heck_10$coefficients)
heck <- data_frame(Coefficient = rn,
                   `Model 10` = heck_10$coefficients,
                   `Model 20` = heck_20$coefficients,
                   `Model 30` = heck_30$coefficients,
                   `Model 40` = heck_40$coefficients,
                   `Model 50` = heck_50$coefficients)
for (j in 2:ncol(heck)){
  heck[,j] <- round(heck[,j], digits = 3)
}
heck$Coefficient <- gsub('bin_citations', 'Citations ', heck$Coefficient)
heck$Coefficient <- gsub('invMillsRatio', 'Inverse Mills Ratio', heck$Coefficient)
heck$Coefficient <- gsub('sigma', 'Sigma', heck$Coefficient)
heck$Coefficient <- gsub('rho', 'Rho', heck$Coefficient)
heck$Coefficient <- gsub('sexmale', 'Sex male', heck$Coefficient)
heck$Coefficient <- paste0(c(rep('Selection ', 4), rep('Outcome ', 2), rep('', 3)), heck$Coefficient)
knitr::kable(heck,
             caption = 'Heckman selection model coefficients')
```

\end{multicols}

\newpage
